Is high social class always beneficial for survival? Northern Sweden 1801--2013

### 2017-11-27 article3.Rmd
affiliation, million

---
title: "Is high social class always beneficial for survival? Northern Sweden 1801--2013"
author: "Sören Edvinsson and Göran Broström"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    citation_package: natbib
    number_sections: yes
    toc: no
    toc_depth: 2
    keep_tex: true
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: yes
    toc_depth: 2
titlepage: yes
bibliography: mybib.bib
documentclass: article
classoption: titlepage
fontsize: 12pt
citation_package: natbib
biblio-style: apalike
header-includes:
  - \usepackage{endfloat}
  - \usepackage[utf8]{inputenc}
abstract: "Focusing on two regions in northern Sweden 1801--2013, we challenge common notions of the assumed advantage in survival of belonging to a high social class. The issue is analysed according to gender and age group (adults and elderly) and in relation to the development of economic inequality.
The results show that high social class is not always favourable for survival. Men in the elite category had higher mortality compared to others during a large part of the studied period; a male mortality class reversal appears at a surprisingly late date, while the social gradient among women conforms to the expected pattern. We suggest that health-related behaviour is decisive not only in later but earlier phases of the mortality transition as well. The results implicate that the association between social class and health is more complex than is assumed in many of the dominant theories in demography and epidemiology."
---

@@

```{r lexis2,fig.cap="The sampling frame.", fig.height = 5}
plot(c(1801, 1801, 1951, 1951, 1801), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1800, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1801, 1851, 1901, 1951, 1976, 1991, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1976, 1976, 2014, 2014, 1976), c(40, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
pn.exp <- round(sum(skel$exposure) + sum(skume$exposure))
pn.d <-  as.integer(sum(skel$event) + sum(skume$event))
#
text(1875, 70, "PopLink", cex = 2, col = "red")
##text(1899, 60, paste(pn.rec, "records"))
text(1875, 55, paste(pn.exp, "p-years"))
text(1874.5, 50, paste(pn.d, "deaths"))
#
n.exp <- round(sum(foli$exposure))
n.d <-  as.integer(sum(foli$event))

text(1995, 70, "Linnaeus", cex = 1.5, col = "blue")
##text(1984, 60, paste(n.rec, "records"))
text(1995.2, 55, paste(n.exp, "p-years"), cex = 0.8)
text(1995.2, 50, paste(n.d, "deaths"))
```

>>>>

---
title: "Is high social class always beneficial for survival? Northern Sweden 1801--2013"
author: Sören Edvinsson\footnote{Corresponding author} and Göran Broström
date: 'Centre for Demographic and Ageing Research \newline Umeå University, SE-90187 Umeå, Sweden'
output:
  bookdown::pdf_document2:
    citation_package: natbib
    number_sections: yes
    toc: no
    toc_depth: 2
    keep_tex: true
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
  bookdown::word_document2:
    toc: yes
    toc_depth: 2
titlepage: yes
bibliography: mybib.bib
documentclass: article
classoption: titlepage
fontsize: 12pt
citation_package: natbib
biblio-style: apalike
header-includes:
  - \usepackage{endfloat}
  - \usepackage[utf8]{inputenc}
abstract: "Focusing on two regions in northern Sweden 1801--2013, we challenge common notions of the assumed advantage in survival of belonging to a high social class. The issue is analysed according to gender and age group (adults and elderly) and in relation to the development of economic inequality.
The results show that high social class is not always favourable for survival. Men in the elite category had higher mortality compared to others during a large part of the studied period; a male mortality class reversal appears at a surprisingly late date, while the social gradient among women conforms to the expected pattern. We suggest that health-related behaviour is decisive not only in later but earlier phases of the mortality transition as well. The results implicate that the association between social class and health is more complex than is assumed in many of the dominant theories in demography and epidemiology."
---

@@

```{r lexis2,fig.cap="The sampling frame.", fig.height = 5}
plot(c(1801, 1801, 1951, 1951, 1801), c(40, 90, 90, 40, 40),
     type = "l", xlim = c(1800, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 90))
axis(1, at = c(1801, 1851, 1901, 1951, 1976, 1991, 2014), las = 2)
axis(2, las = 2, at = c(0, 40, 65, 90))
axis(4, las = 2, at = c(0, 40, 65, 90))
box()
abline(h = 0)
lines(c(1976, 1976, 2014, 2014, 1976), c(40, 90, 90, 40, 40),
      col = "blue", lwd = 2)
#
pn.exp <- round(sum(skel$exposure) + sum(skume$exposure))
pn.d <-  as.integer(sum(skel$event) + sum(skume$event))
#
text(1875, 70, "PopLink", cex = 2, col = "red")
##text(1899, 60, paste(pn.rec, "records"))
text(1875, 55, paste(round(pn.exp / 1000000, 1), "million p-years"))
text(1874.5, 50, paste(round(pn.d / 1000, 1), "thousand deaths"))
#
n.exp <- round(sum(foli$exposure))
n.d <-  as.integer(sum(foli$event))

text(1995, 70, "Linnaeus", cex = 1.5, col = "blue")
##text(1984, 60, paste(n.rec, "records"))
text(1995.2, 55, paste(round(n.exp / 1000000, 1), "million p-years"), cex = 0.8)
text(1995.2, 50, paste(round(n.d / 1000, 1), "thousand deaths"), cex = 0.7)
```
